### 2022-01-21 Telling-stories-with-data.Rmd
Create paper 1 project

---
title: "Telling stories"
author: "Alicia Yang"
date: "09/01/2022"
output: pdf_document
abstract: "This paper"
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2022-02-02 01-data_cleaning.R
load data

@@ -1,0 +1,75 @@
+#### Preamble ####
+# Purpose: Clean the survey data downloaded from [...UPDATE ME!!!!!]
+# Author: Rohan Alexander [CHANGE THIS TO YOUR NAME!!!!]
+# Data: 3 January 2021
+# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
+# License: MIT
+# Pre-requisites: 
+# - Need to have downloaded the ACS data and saved it to inputs/data
+# - Don't forget to gitignore it!
+# - Change these to yours
+# Any other information needed?
+
+
+#### Workspace setup ####
+# Use R Projects, not setwd().
+library(haven)
+library(tidyverse)
+library(opendatatoronto)
+library(dplyr)
+
+
+# get package
+package <- show_package("0b6d3a00-7de1-440b-b47c-7252fd13929f")
+package
+
+# get all resources for this package
+resources <- list_package_resources("0b6d3a00-7de1-440b-b47c-7252fd13929f")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+data1 <-
+  list_package_resources("0b6d3a00-7de1-440b-b47c-7252fd13929f") %>%
+  get_resource()
+
+# get package
+package2 <- show_package("9fcff3e1-3737-43cf-b410-05acd615e27b")
+package2
+
+# get all resources for this package
+resources2 <- list_package_resources("9fcff3e1-3737-43cf-b410-05acd615e27b")
+
+data2 <-
+  list_package_resources("9fcff3e1-3737-43cf-b410-05acd615e27b") %>%
+  get_resource()
+
+write_csv(
+  x = data1,
+  file = "Telling-stories-with-data/inputs/data/vehicle_collisions.csv"
+)
+
+write_csv(
+  x = data2,
+  file = "Telling-stories-with-data/inputs/data/red_light_cameras.csv"
+)
+
+# Read in the raw data. 
+raw_data <- readr::read_csv("inputs/data/raw_data.csv"
+                     )
+# Just keep some variables that may be of interest (change 
+# this depending on your interests)
+names(raw_data)
+
+reduced_data <- 
+  raw_data %>% 
+  select(first_col, 
+         second_col)
+rm(raw_data)
+         
+
+#### What's next? ####
+
+
+
+         

### 2022-02-03 outputs/paper/paper.Rmd
introduction

---
title: "Toronto motor vehicle collisions can be reduced by installing more red cameras"
subtitle: "Toronto motor vehicle collisions can be reduced by installing more red cameras"
author: 
  - Alicia Yang
thanks: "Code and data are available at: https://github.com/Alicia-y/Telling-stories-with-data."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Vehicle collisions occur constantly in big citys. They sometimes could be severe issues when resulting in damage in public and personal properties, injuries, or even death. This report aims to investigate the motor vehicle collisions in Toronto that result in injuries or deaths from 2006 to 2020. By exploring the trend of these traffic collisions in respect of number and location, I find that .... Implies that ... The data used in this report is obtained from the City of Toronto Open Data Portal, and ."
output:
  bookdown::pdf_document2
toc: TRUE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(palmerpenguins)
```

\newpage

# Introduction

As one of the busiest and most modern city in Canada, Toronto constantly faces a large number of citizens, and, inevitably, heavy traffic. The road condition in Toronto is busy most of the time. The vast majority of roads and buildings under construction that take place almost everywhere narrow the roadway of vehicles and turn off traffic lights. Heavy traffic is definitely one of the main reasons behind car accidents, because it increases road rage and encourage the vehicle drivers to make dangerous moves to save time. Vehicle collisions have always brought large financial costs to the involving parties, and even more, they even brought injuries and deaths.

The Motor Vehicle Collisions data of Toronto provides a record of the past severe vehicle collisions in Toronto, and with the help of statistical data analysis, it can possibly offer insights on the ways to reduce this type of accident. This report investigates the motor vehicle collisions in Toronto from 2006 to 2020, with a focus on the ones result in injuries or deaths. In Section \@ref(data), the Motor Vehicle Collisions dataset is introduced and organized to visualize and explain the change in the annual number of accidents, and the frequency of the accidents by district and by type.

However, numbers of annual motor vehicle collisions demonstrate a generally descending trend from 2006 to 2020.

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data
To investigate the motor vehicle collisions in Toronto, I obtained the dataset "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" from the Toronto Open Data Portal. Provided by the Toronto Police Service, this dataset included all vehicle collisions occurred in Toronto between 2006 and 2020 that result in injuries or deaths.

```{r, echo=FALSE}
vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
  clean_names()
```

Show an extract of the cleaned dataset (Table \@ref(tab:dataextract))

```{r dataextract}
library(knitr)
vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
  clean_names()
vehicle_collision %>%
  select(year, street1, street2, accloc, neighbourhood) %>%
  slice(1:10) %>%
  kable(
    caption = "Extracting the first ten rows from the Motor Vehicle Collisions data",
col.names=c("Year", "Street1", "Street2", "Collision location", "Neighbourhood"),
linesep = "", digits = 1, booktabs=TRUE)

```
Table \@ref(tab:dataextract) shows

```{r fig.cap = "Comparing the number of traffic collisions each year from 2006 to 2020"}
ggplot(mapping = aes(x = year), data=vehicle_collision) +
  geom_bar()
```

Data cleaning
```{r}
vehicle_collision <- vehicle_collision %>%
  drop_na(street1) %>%
  drop_na(street2) %>%
  drop_na(accloc) %>%
  drop_na(neighbourhood)
```

```{r fig.cap = "Comparing the location of traffic collisions each year from 2006 to 2020"}
ggplot(mapping = aes(x = year, fill=accloc), data=vehicle_collision) +
  geom_bar()
```

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

Yihui Xie (2021). knitr: A General-Purpose Package for Dynamic Report Generation in R. R
  package version 1.37.

https://newcanadianlife.com/why-does-toronto-have-so-much-traffic/

### 2022-02-03 outputs/paper/paper.Rmd
data1

@@
 The vast majority of roads and buildings under construction that take place almost everywhere narrow the roadway of vehicles and turn off traffic lights.
-Heavy traffic is definitely one of the main reasons behind car accidents, because it increases road rage and encourage the vehicle drivers to make dangerous moves to save time.
+Heavy traffic is likely to be one of the main reasons behind car accidents, because it increases the road rage and encourages the vehicle drivers to make dangerous moves to save time.
 Vehicle collisions have always brought large financial costs to the involving parties, and even more, they even brought injuries and deaths.
 
-The Motor Vehicle Collisions data of Toronto provides a record of the past severe vehicle collisions in Toronto, and with the help of statistical data analysis, it can possibly offer insights on the ways to reduce this type of accident.
+The Motor Vehicle Collisions data of Toronto provides a record of the past severe vehicle collisions in Toronto which result in injuries or deaths, and with the help of statistical data analysis, it can possibly offer insights on the ways to reduce this type of accident.
 This report investigates the motor vehicle collisions in Toronto from 2006 to 2020, with a focus on the ones result in injuries or deaths.
-In Section \@ref(data), the Motor Vehicle Collisions dataset is introduced and organized to visualize and explain the change in the annual number of accidents, and the frequency of the accidents by district and by type.
+In Section \@ref(data), the Motor Vehicle Collisions dataset is introduced and organized to visualize the change in the annual number of accidents, as well as comparing the frequency of the accidents by district, accident type, and level of injury.
+Then, I discussed about the possible implications of these visualizations.
 
@@
 To investigate the motor vehicle collisions in Toronto, I obtained the dataset "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" from the Toronto Open Data Portal.
-Provided by the Toronto Police Service, this dataset included all vehicle collisions occurred in Toronto between 2006 and 2020 that result in injuries or deaths.
+Provided by the Toronto Police Service, this dataset includes all vehicle collisions occurred in Toronto between 2006 and 2020 that result in injuries or deaths.
+It contains 16840 observations of collisions and 54 variables with all kinds of information of each collision.
+This report focuses on six of these variables which are: year, accident location, district, street1(of the collision occurred), street2(of the collision occurred), and classification of accident.
+By using R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" (Yihui Xie 2021), an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.
 
-```{r, echo=FALSE}
-vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
-  clean_names()
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+vehicle_collision <- read_csv(here::here("Telling-stories-with-data/inputs/data/vehicle_collisions.csv")) %>%
+  clean_names() %>%
+  drop_na(year)
 ```
 
-Show an extract of the cleaned dataset (Table \@ref(tab:dataextract))
-
-```{r dataextract}
+```{r dataextract, echo=FALSE, message=FALSE}
 library(knitr)
-vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
-  clean_names()
+library(kableExtra)
+vehicle_collision <- vehicle_collision %>%
+  select(year, street1, street2, accloc, district, acclass)
 vehicle_collision %>%
-  select(year, street1, street2, accloc, neighbourhood) %>%
   slice(1:10) %>%
@@
     caption = "Extracting the first ten rows from the Motor Vehicle Collisions data",
-col.names=c("Year", "Street1", "Street2", "Collision location", "Neighbourhood"),
-linesep = "", digits = 1, booktabs=TRUE)
+col.names=c("Year", "Street1", "Street2", "Collision location", "District", "Classification of Accident
+
+"),
+linesep = "", digits = 1, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position")
 
 ```
-Table \@ref(tab:dataextract) shows
 
-```{r fig.cap = "Comparing the number of traffic collisions each year from 2006 to 2020"}
-ggplot(mapping = aes(x = year), data=vehicle_collision) +
-  geom_bar()
+Table \@ref(tab:dataextract) shows the first ten rows of the Motor Vehicle Collisions data.
+Variable "Year" indicates the year in which the accident occurs.
+Variables "Street1" and "Street2" provides the name of the street where the accident takes place.
+Variable "District" indicates its district.
+Variable "Collision location" includes 8 types: at intersection, intersection related, at/near private drive, laneway, non intersection, overpass or bridge, trail, and underpass or tunnel.
+Variable "Classification of Accident" involves fatal and non-fatal injury.
+
+I'm interested in finding the trend of motor vehicle collisions from 2006 to 2020 and the types of injuries they result in.
+It reveals the traffic control in Toronto, and possibly the quality and knowledge of vehicle owners throughout these years.
+
+```{r fig.cap = "Comparing the number of traffic collisions each year from 2006 to 2020", fig.width=5.4,fig.height=3.6, echo=FALSE}
+ggplot(mapping = aes(x = year, fill = acclass), data=vehicle_collision) +
+  geom_bar()+theme_light()+ylab("number of motor vehicle collisions involving injuries or deaths")+scale_fill_discrete(name = "Classification of Accident")
 ```
-
-Data cleaning
-```{r}
-vehicle_collision <- vehicle_collision %>%
-  drop_na(street1) %>%
-  drop_na(street2) %>%
-  drop_na(accloc) %>%
-  drop_na(neighbourhood)
-```
+Figure 1 shows a descending pattern for the number of car accidents involving injuries or deaths as the number went from 1483 cases in 2006 to 633 cases in 2020.
+By looking at the accident types,  Indicate that it's likely that Toronto has done a better job on traffic control in general.
+However, if other factors are taken into considerations, the trend of motor
 

### 2022-02-04 outputs/paper/paper.Rmd
data2

@@
 ---
-title: "Toronto motor vehicle collisions can be reduced by installing more red cameras"
-subtitle: "Toronto motor vehicle collisions can be reduced by installing more red cameras"
+title: "Investigating the motor vehicle collisions in Toronto"
+subtitle: "Installing more red light cameras do not stop traffic accidents from happening"
 author: 
@@
 This report aims to investigate the motor vehicle collisions in Toronto that result in injuries or deaths from 2006 to 2020.
-By exploring the trend of these traffic collisions in respect of number and location, I find that ....
-Implies that ...
-The data used in this report is obtained from the City of Toronto Open Data Portal, and ."
+By exploring the trend of these traffic collisions in respect of number and location, I find that the total number of accidents has a descending trend.
+However, the vehicle collisions taking place at intersections has not been reduced despite that more red light cameras are installed.
+The data used in this report is obtained from the City of Toronto Open Data Portal."
 output:
@@
 Vehicle collisions have always brought large financial costs to the involving parties, and even more, they even brought injuries and deaths.
+In just 2019, traffic accidents resulted in more than 1700 deaths and more than 140000 injuries.
 
@@
 # Data
+## Motor Vehicle Collisions involving Killed or Seriously Injured Persons
 To investigate the motor vehicle collisions in Toronto, I obtained the dataset "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" from the Toronto Open Data Portal.
@@
 This report focuses on six of these variables which are: year, accident location, district, street1(of the collision occurred), street2(of the collision occurred), and classification of accident.
-By using R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" (Yihui Xie 2021), an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.
+By using R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr], an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.
 
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-vehicle_collision <- read_csv(here::here("Telling-stories-with-data/inputs/data/vehicle_collisions.csv")) %>%
+vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
   clean_names() %>%
-  drop_na(year)
+  drop_na(year) %>%
+  filter(acclass != "Property Damage Only")
 ```
@@
 linesep = "", digits = 1, booktabs=TRUE) %>%
-  kable_styling(latex_options = "HOLD_position")
+  kable_styling(latex_options = "HOLD_position", font_size = 7)
 
@@
 
-```{r fig.cap = "Comparing the number of traffic collisions each year from 2006 to 2020", fig.width=5.4,fig.height=3.6, echo=FALSE}
+```{r fig.cap = "Comparing the number of traffic collisions each year from 2006 to 2020", fig.width=8,fig.height=3.4, echo=FALSE}
 ggplot(mapping = aes(x = year, fill = acclass), data=vehicle_collision) +
-  geom_bar()+theme_light()+ylab("number of motor vehicle collisions involving injuries or deaths")+scale_fill_discrete(name = "Classification of Accident")
+  geom_bar()+theme_light()+ylab("number of motor vehicle collisions")+scale_fill_discrete(name = "Classification of Accident")+facet_wrap(~acclass)
 ```
-Figure 1 shows a descending pattern for the number of car accidents involving injuries or deaths as the number went from 1483 cases in 2006 to 633 cases in 2020.
-By looking at the accident types,  Indicate that it's likely that Toronto has done a better job on traffic control in general.
-However, if other factors are taken into considerations, the trend of motor
 
-```{r fig.cap = "Comparing the location of traffic collisions each year from 2006 to 2020"}
-ggplot(mapping = aes(x = year, fill=accloc), data=vehicle_collision) +
+Figure 1 shows a descending pattern for the number of car accidents involving non-fatal injuries from year 2006 to 2020.
+The number of fatal accidents does not increase nor decrease much.
+In general, the number vehicle collisions in Toronto has largely decrease, indicating that when combining this two factors(number and classification of accident), Toronto has done a better job on traffic control in general.
+I'm also interested in the trend of motor vehicle collisions if other factors are taken into considerations.
+\newpage
+
+```{r fig.cap = "Comparing the location of traffic collisions each year from 2006 to 2020", echo=FALSE, fig.width=6,fig.height=3.4}
+ggplot(mapping = aes(x = year, fill=accloc), data=vehicle_collision)+scale_fill_discrete(name = "Accident location")+ylab("number of motor vehicle collisions")+theme_light()+
   geom_bar()
+```
+
+Figure 2 illustrates the where these vehicle collisions usually take place from 2006 to 2020.
+Before 2013, half of the data doesn't come with accident location, which is understandable because back in the days there wasn't an easy approach to record and save the data, so they often got lost.
+A large proportion of the accidents take place either at intersections or is intersection related.
+The other seven accident locations types add up to less than half of all vehicle collisions.
+At intersections, a lot more rule breaking behaviors can be committed by drivers such as ignoring stop signs or traffic lights, some drivers even claimed that they rarely yielded.
+These behaviors would result in collisions with other vehicles and the pedestrians.
+
+Motor vehicle collisions are most likely to take place at intersections.
+If we could measures are taken to restrict the driver's rule breaking behaviors at intersections, the numbers of car accidents would drop a lot in general.
+Red light camera is supposed to be a measure taken by the government to monitor the vehicles at intersections.
+I will look into whether it's effective or not in respect of preventing traffic accidents from happening.
+
+## Red Light Cameras
+The number of red light camera instilled in Toronto from 2007 to 2020 is obtained from the dataset "Red Light Cameras" found in the Toronto Open Data Portal.
+This dataset is shared by the Transportation Services and it covers the locations of red light cameras in Toronto.
+In its introduction, this dataset claims that red light cameras can improve the safety of the drivers and pedestrians by monitoring the behavior of the drivers.
+```{r message=FALSE, warning=FALSE}
+red_light_camera <- read_csv(here::here("inputs/data/red_light_cameras.csv")) %>%
+  clean_names() %>%
+  select(activation_date, district) %>%
+  drop_na(activation_date) %>%
+  drop_na(district)
+red_light_camera$year <- vector("numeric", length = 199)
+for (i in 1:199){
+  red_light_camera$year[i]=format(as.Date(red_light_camera$activation_date[i], format="%Y/%m/%d"),"%Y")
+}
+red_light_camera <- red_light_camera %>%
+  filter(year != 2021)
+cumulative <-tibble(
+  year=c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020),
+  sum=c(0, 2, 6, 15, 16, 17, 17, 17, 17, 17, 17, 31, 33, 33, 34, 0, 0, 7, 17, 18, 18, 18, 18, 18, 18, 18, 42, 43, 44, 44, 0, 3, 10, 18, 19, 19, 19, 19, 19, 19, 19, 26, 27, 28, 28, 0, 1, 9, 15, 15, 15, 15, 15, 15, 15, 15, 36, 38, 41, 41),
+  district=c("Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York")
+)
+ggplot( aes(x=year, y=sum, group=district, color=district), data=cumulative) +
+    geom_line()
+```
+
+```{r}
+co <- vehicle_collision %>%
+  filter(accloc == "At Intersection" | accloc == "Intersection Related") %>%
+  drop_na(district)
+for(i in 1:9039){
+  if(co$district[i] == "Toronto East York"){
+    co$district[i] = "Toronto and East York"
+  }
+}
+ggplot(data=co,aes(x=year, fill=acclass))+geom_bar()+facet_wrap(~district)
 ```
@@
 
-Yihui Xie (2021). knitr: A General-Purpose Package for Dynamic Report Generation in R. R
-  package version 1.37.
+https://tc.canada.ca/en/road-transportation/statistics-data/canadian-motor-vehicle-traffic-collision-statistics-2019
 

### 2022-02-05 outputs/paper/paper.Rmd
finish data

--- 2022-02-03 outputs/paper/paper.Rmd
@@
 ---
-title: "Toronto motor vehicle collisions can be reduced by installing more red cameras"
-subtitle: "Toronto motor vehicle collisions can be reduced by installing more red cameras"
+title: "Investigating motor vehicle collisions in Toronto: Installing more red light cameras do not stop traffic accidents from happening"
 author: 
@@
 They sometimes could be severe issues when resulting in damage in public and personal properties, injuries, or even death.
-This report aims to investigate the motor vehicle collisions in Toronto that result in injuries or deaths from 2006 to 2020.
-By exploring the trend of these traffic collisions in respect of number and location, I find that ....
-Implies that ...
-The data used in this report is obtained from the City of Toronto Open Data Portal, and ."
+This report aims to investigate motor vehicle collisions in Toronto that result in injuries or deaths from 2006 to 2020.
+By exploring the trend of these traffic collisions in respect of their frequency, fatality, and location, I find that the total number of accidents has a descending trend.
+However, the number of vehicle collisions that occurred at intersections has not reduced despite the fact that more red light cameras are installed.
+This finding suggests that ...
+furthur measures should be taken to enhance the safety issue on road."
 output:
@@
 knitr::opts_chunk$set(echo = TRUE)
-
+# Set up libraries
 library(tidyverse)
 library(janitor)
-library(palmerpenguins)
 ```
@@
 This report investigates the motor vehicle collisions in Toronto from 2006 to 2020, with a focus on the ones result in injuries or deaths.
-In Section \@ref(data), the Motor Vehicle Collisions dataset is introduced and organized to visualize the change in the annual number of accidents, as well as comparing the frequency of the accidents by district, accident type, and level of injury.
+In Section \@ref(data), the Motor Vehicle Collisions dataset is introduced and organized to visualize the change in the annual number of traffic accidents with respect to their fatality and their location.
 Then, I discussed about the possible implications of these visualizations.
-
-However, numbers of annual motor vehicle collisions demonstrate a generally descending trend from 2006 to 2020.
+By Figure 2, I concluded that motor vehicle collisions are always more likely to occur at intersections.
+Thus,
+Continuing
 
@@
 # Data
-To investigate the motor vehicle collisions in Toronto, I obtained the dataset "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" from the Toronto Open Data Portal.
+## Motor Vehicle Collisions involving Killed or Seriously Injured Persons
+To investigate the motor vehicle collisions in Toronto, I obtained the dataset "Motor Vehicle Collisions involving Killed or Seriously Injured Persons" from the Toronto Open Data Portal [@citedata].
 Provided by the Toronto Police Service, this dataset includes all vehicle collisions occurred in Toronto between 2006 and 2020 that result in injuries or deaths.
-It contains 16840 observations of collisions and 54 variables with all kinds of information of each collision.
-This report focuses on six of these variables which are: year, accident location, district, street1(of the collision occurred), street2(of the collision occurred), and classification of accident.
-By using R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" (Yihui Xie 2021), an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.
+One bias aroused by this dataset is the offset from the real accident location.
+For ethical reason and the protection of the parties involved in the accidents that are considered to be criminal cases, the location of those criminal accidents are offset to the closest intersection node [@citecollisiondata].
+As a result, the location of those accidents specific to neighborhoods might not be accurate.
+Therefore, I will use the variable district instead of neighbourhood to categorize the accidents later in the report to avoid this bias.
+
+This dataset contains 16840 observations of collisions and 54 variables with all kinds of information of each collision.
+This report focuses on four of these variables which are: year, accident location, district, and fatality.
+By using R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr], an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.
 
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-vehicle_collision <- read_csv(here::here("Telling-stories-with-data/inputs/data/vehicle_collisions.csv")) %>%
+# Load motor vehicle collision data and clean
+vehicle_collision <- read_csv(here::here("inputs/data/vehicle_collisions.csv")) %>%
   clean_names() %>%
-  drop_na(year)
+  drop_na(year) %>%
+  filter(acclass != "Property Damage Only")
 ```
@@
 ```{r dataextract, echo=FALSE, message=FALSE}
+# Set up libraries
 library(knitr)
 library(kableExtra)
+
+# Create a table with selected columns
 vehicle_collision <- vehicle_collision %>%
-  select(year, street1, street2, accloc, district, acclass)
+  select(year, accloc, district, acclass)
 vehicle_collision %>%
@@
     caption = "Extracting the first ten rows from the Motor Vehicle Collisions data",
-col.names=c("Year", "Street1", "Street2", "Collision location", "District", "Classification of Accident
-
-"),
+col.names=c("Year", "Collision location", "District", "Fatality"),
 linesep = "", digits = 1, booktabs=TRUE) %>%
-  kable_styling(latex_options = "HOLD_position")
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
 
@@
 Variable "Year" indicates the year in which the accident occurs.
-Variables "Street1" and "Street2" provides the name of the street where the accident takes place.
-Variable "District" indicates its district.
+Variable "District" provides the district of the accident.
 Variable "Collision location" includes 8 types: at intersection, intersection related, at/near private drive, laneway, non intersection, overpass or bridge, trail, and underpass or tunnel.
-Variable "Classification of Accident" involves fatal and non-fatal injury.
+Variable "Fatality" involves fatal and non-fatal injury.
 
@@
 
-```{r fig.cap = "Comparing the number of traffic collisions each year from 2006 to 2020", fig.width=5.4,fig.height=3.6, echo=FALSE}
+```{r fig.cap = "Comparing the number of traffic collisions each year from 2006 to 2020", fig.width=8.4,fig.height=3.6, echo=FALSE}
+# Create a bar plot to show number of collisions by fatality
 ggplot(mapping = aes(x = year, fill = acclass), data=vehicle_collision) +
-  geom_bar()+theme_light()+ylab("number of motor vehicle collisions involving injuries or deaths")+scale_fill_discrete(name = "Classification of Accident")
-```
-Figure 1 shows a descending pattern for the number of car accidents involving injuries or deaths as the number went from 1483 cases in 2006 to 633 cases in 2020.
-By looking at the accident types,  Indicate that it's likely that Toronto has done a better job on traffic control in general.
-However, if other factors are taken into considerations, the trend of motor
-
-```{r fig.cap = "Comparing the location of traffic collisions each year from 2006 to 2020"}
-ggplot(mapping = aes(x = year, fill=accloc), data=vehicle_collision) +
-  geom_bar()
+  geom_bar()+theme_light()+ylab("number of motor vehicle collisions")+scale_fill_discrete(name = "Fatality")+facet_wrap(~acclass)
 ```
 
-```{r bills, fig.cap="Bills of penguins", echo = FALSE}
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+Figure 1 shows a descending pattern for the number of car accidents involving non-fatal injuries from year 2006 to 2020.
+The number of fatal accidents does not increase nor decrease much.
+In general, the number vehicle collisions in Toronto has largely decrease, indicating that when combining this two factors(number and fatality), Toronto has done a better job on traffic control in general.
+I'm also interested in the trend of motor vehicle collisions if other factors are taken into considerations.
+\newpage
+
+```{r fig.cap = "Comparing the location of traffic collisions each year from 2006 to 2020", echo=FALSE, fig.width=6,fig.height=3.2}
+# Create a bar plot to show number of collisions by accident location
+ggplot(mapping = aes(x = year, fill=accloc), data=vehicle_collision)+scale_fill_discrete(name = "Accident location")+ylab("number of motor vehicle collisions")+theme_light()+geom_bar()
 ```
 
-Talk more about it.
+Figure 2 illustrates the where these vehicle collisions usually take place from 2006 to 2020.
+Before 2013, half of the data doesn't come with accident location, which is understandable because back in the days there wasn't an easy approach to record and save the data, so they often got lost.
+A large proportion of the accidents take place either at intersections or is intersection related.
+The other seven accident locations types add up to less than half of all vehicle collisions.
+At intersections, a lot more rule breaking behaviors can be committed by drivers.
+For instance, they could sometimes ignore stop signs, speed up when approaching yellow traffic lights; some even claimed that they rarely yielded at intersections.
+These behaviors largely increase the possibility of them colliding with other vehicles and the pedestrians at intersections.
 
-Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)
+Motor vehicle collisions are most likely to take place at intersections.
+Thus, if measures are taken to restrict the driver's rule breaking behaviors at intersections, the numbers of car accidents would certainly drop a lot in general.
+Red light camera is claimed to be a measure taken by the government to monitor the vehicles at intersections.
+I will look into whether it's effective or not in respect of preventing the traffic accidents from happening.
 
-```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}
+## Red Light Cameras
+The number of red light camera instilled in Toronto from 2007 to 2020 is obtained from the dataset "Red Light Cameras" found in the Toronto Open Data Portal [@citedata].
+This dataset is shared by the Transportation Services and it covers the locations of red light cameras in Toronto.
+In its introduction, this dataset claims that red light cameras can improve the safety of the drivers and pedestrians by monitoring the behavior of the drivers [@citecameradata].
+It contains 199 observations and 32 variables.
+This dataset's downside is that it is missing values in lots of its variables.
+To reduce errors generated from this downside, I will only utilize the variables with no missing values.
+R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr] are usedt to produce an extract of the cleaned dataset (Table \@ref(tab:dadaextract1)).
 
-# This needs to be about the random data tha tI sasmpled?
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+```{r dadaextract1, message=FALSE, warning=FALSE, echo=FALSE}
+# Load red light camera data and clean
+red_light_camera <- read_csv(here::here("inputs/data/red_light_camera.csv")) %>%
+  clean_names() %>%
+  select(activation_date, district) %>%
+  drop_na(activation_date) %>%
+  drop_na(district)
+
+# Create "year" variable
+red_light_camera$year <- vector("numeric", length = 199)
+for (i in 1:199){
+  red_light_camera$year[i]=format(as.Date(red_light_camera$activation_date[i], format="%Y/%m/%d"),"%Y")
+}
+
+# Create a table with selected columns
+red_light_camera %>%
+  select(district, year, activation_date) %>%
+  slice(1:10) %>%
+  kable(
+    caption = "Extracting the first ten rows from the Red Light Camera data",
+col.names=c("District", "Year", "Activation date"),
+linesep = "", digits = 1, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
 ```
 
-Talk way more about it.
+Table \@ref(tab:dadaextract1) shows the first ten rows of the Red Light Camera data in Toronto.
+"Year" is a new variable added by me into this dataset.
+It is pulled from the original variable "activation date", and it provides the year in which the red light camera is activated.
+Variable "District" includes the four main districts of Toronto: Toronto and East York, North York, Scarborough, and Etobicoke York.
 
-# Model
+I'm interested in observing the cumulative number of red light cameras installed in each district from 2006 to 2020 and compare this data with the trend of number of motor vehicle collisions to see if they actually have connections.
 
-\begin{equation}
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
-\end{equation}
+```{r message=FALSE, warning=FALSE, echo=FALSE, fig.cap = "Number of red light cameras in the four districts each year", fig.width=7,fig.height=3.2}
+# Create a new dataset that presents the total number of red light cameras in each district
+red_light_camera <- red_light_camera %>%
+  filter(year != 2021)
+cumulative <-tibble(
+  year=c(2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020),
+  sum=c(0, 2, 6, 15, 16, 17, 17, 17, 17, 17, 17, 31, 33, 33, 34, 0, 0, 7, 17, 18, 18, 18, 18, 18, 18, 18, 42, 43, 44, 44, 0, 3, 10, 18, 19, 19, 19, 19, 19, 19, 19, 26, 27, 28, 28, 0, 1, 9, 15, 15, 15, 15, 15, 15, 15, 15, 36, 38, 41, 41),
+  district=c("Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Toronto and East York", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "Scarborough", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "North York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York", "Etobicoke York")
+)
 
-Equation \@ref(eq:bayes) seems useful, eh?
+#Create a line plot from the dataset
+ggplot( aes(x=year, y=sum, group=district, color=district), data=cumulative) +
+    geom_line()+ylab("number of red light cameras")+theme_light()
+```
 
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
+Figure 3 demonstrates the total number of red light cameras in each district from 2006 to 2020.
+Scarborough installed the most red light cameras in this period of time, especially after 2016, a large number of red light cameras were installed there.
+The district of North York, on the contrary, started with the most red light cameras installed among all districts, but has the slowest growth and ended up with the least number of red light cameras in 2020 among all districts.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+I wanted to see how the increase in number of red light cameras compare with the traffic collision's frequency in each district.
+So, I selected the accidents that occurred at intersections or somewhere near intersections from the Motor Vehicle Collision dataset and created a ggplot to see their trend by each district in Toronto.
 
-# Results
+```{r, echo=FALSE, warning=FALSE, fig.cap = "Comparing the number of motor vehicle collisions take place at or near intersections by district", fig.width=7,fig.height=3.2}
+# Select the accidents that only occur at or near intersections
+co <- vehicle_collision %>%
+  filter(accloc == "At Intersection" | accloc == "Intersection Related") %>%
+  drop_na(district)
+for(i in 1:9039){
+  if(co$district[i] == "Toronto East York"){
+    co$district[i] = "Toronto and East York"
+  }
+}
+
+# Create bar plot to compare vehicle collisions by districts
+ggplot(data=co,aes(x=year, fill=acclass))+geom_bar()+facet_wrap(~district)+ylab("number of motor vehicle collisions")+theme_light()+scale_fill_discrete(name = "Fatality")
+```
+
+Figure 4 shows the number of annual motor vehicle accidents occurred at or near intersections in each district.
+For all the districts, the numbers of fatal vehicle collisions stay approximately the same throughout these years, with Scarborough having the highest mortality rate in general.
+With regards to the total number of accidents, data in Etobicoke York, North York, and Scarborough all has its fluctuations, but the number remains about the same.
+The number of vehicle collisions in Toronto and East York was the highest in 2006, with more than to 260 cases, but it is the only district where the number has dropped.
+
+By comparing Figure 3 and Figure 4, I found that they haven't show any associations between the number of traffic accidents and the number of red light cameras installed.
 
@@
 
-Yihui Xie (2021). knitr: A General-Purpose Package for Dynamic Report Generation in R. R
-  package version 1.37.
+https://tc.canada.ca/en/road-transportation/statistics-data/canadian-motor-vehicle-traffic-collision-statistics-2019
 
 https://newcanadianlife.com/why-does-toronto-have-so-much-traffic/
+
+https://www.scientificamerican.com/article/red-light-cameras-may-not-make-streets-safer/

### 2022-02-05 01-data_cleaning.R

@@ -21,6 +21,7 @@ library(dplyr)
 
 # get package
 package <- show_package("0b6d3a00-7de1-440b-b47c-7252fd13929f")
+
 package
 
 # get all resources for this package
@@ -51,25 +52,8 @@ write_csv(
 
 write_csv(
   x = data2,
-  file = "Telling-stories-with-data/inputs/data/red_light_cameras.csv"
+  file = "Telling-stories-with-data/inputs/data/red_light_camera.csv"
 )
-
-# Read in the raw data. 
-raw_data <- readr::read_csv("inputs/data/raw_data.csv"
-                     )
-# Just keep some variables that may be of interest (change 
-# this depending on your interests)
-names(raw_data)
-
-reduced_data <- 
-  raw_data %>% 
-  select(first_col, 
-         second_col)
-rm(raw_data)
-         
-
-#### What's next? ####
-
 
 
          

### 2022-02-05 outputs/paper/paper.Rmd
finalize data

@@
 date: "`r format(Sys.time(), '%d %B %Y')`"
-abstract: "Vehicle collisions occur constantly in big citys.
-They sometimes could be severe issues when resulting in damage in public and personal properties, injuries, or even death.
+abstract: "Vehicle collisions occur constantly in Toronto and sometimes could be severe when resulting in damage in public and personal properties, injuries, or even death.
 This report aims to investigate motor vehicle collisions in Toronto that result in injuries or deaths from 2006 to 2020.
@@
 However, the number of vehicle collisions that occurred at intersections has not reduced despite the fact that more red light cameras are installed.
-This finding suggests that ...
-furthur measures should be taken to enhance the safety issue on road."
+This finding suggests that unlike, ......
+doesn't prevent traffic accidents at intersection, furthur measures should be taken to enhance the safety issue on road."
 output:
@@
 Vehicle collisions have always brought large financial costs to the involving parties, and even more, they even brought injuries and deaths.
-In just 2019, traffic accidents resulted in more than 1700 deaths and more than 140000 injuries.
+In just 2019, traffic accidents resulted in more than 1700 deaths and more than 140000 injuries [@cite1].
 
@@
 By comparing Figure 3 and Figure 4, I found that they haven't show any associations between the number of traffic accidents and the number of red light cameras installed.
+For instance, Scarborough has installed the most number of red light cameras, especially after the year of 2016, but its number of traffic accidents didn't reduced at all, it even reach its peak in 2018, nor did the fatality of those accidents.
+We see a large decrease in Toronto and East York's motor vehicle collisions, but it didn't installed at much red cameras as Etobicoke York or Scarborough.
 
@@
 
-## First discussion point
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
+No relationships
+No increase non decrease
 
@@
 
-## Weaknesses and next steps
+## Limitations
 
@@
 
-https://tc.canada.ca/en/road-transportation/statistics-data/canadian-motor-vehicle-traffic-collision-statistics-2019
-
 https://newcanadianlife.com/why-does-toronto-have-so-much-traffic/

### 2022-02-06 outputs/paper/paper.Rmd
discussion

@@
 By exploring the trend of these traffic collisions in respect of their frequency, fatality, and location, I find that the total number of accidents has a descending trend.
-However, the number of vehicle collisions that occurred at intersections has not reduced despite the fact that more red light cameras are installed.
-This finding suggests that unlike, ......
-doesn't prevent traffic accidents at intersection, furthur measures should be taken to enhance the safety issue on road."
+However, the number of vehicle collisions that occurred at intersections has not been reduced despite the fact that more red light cameras are installed.
+This finding suggests that unlike the claimed purpose of this electronic monitor, red light camera doesn't prevent serious traffic accidents from happening at intersection, and furthur measures should be taken to enhance the safety issue on road."
 output:
@@
 The road condition in Toronto is busy most of the time.
-The vast majority of roads and buildings under construction that take place almost everywhere narrow the roadway of vehicles and turn off traffic lights.
+The vast majority of roads and buildings under construction that take place almost everywhere narrow the roadway of vehicles and turn off traffic lights [@cite4].
 Heavy traffic is likely to be one of the main reasons behind car accidents, because it increases the road rage and encourages the vehicle drivers to make dangerous moves to save time.
@@
 This report focuses on four of these variables which are: year, accident location, district, and fatality.
-By using R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr], an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.
+By using R [@citeR], and R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr], an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.
 
@@
 To reduce errors generated from this downside, I will only utilize the variables with no missing values.
-R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr] are usedt to produce an extract of the cleaned dataset (Table \@ref(tab:dadaextract1)).
+R [@citeR], R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr] are used to produce an extract of the cleaned dataset (Table \@ref(tab:dadaextract1)).
 
@@
 The district of North York, on the contrary, started with the most red light cameras installed among all districts, but has the slowest growth and ended up with the least number of red light cameras in 2020 among all districts.
+All the four districts installed a large number of red light cameras from 2016 to 2017.
 
@@
 # Discussion
+The analysis of the two datasets show no relationship between the cases of motor vehicle collisions in Toronto and the number of red light cameras.
+From the comparision between Figure 3 and Figure 4, we see that in none of the districts did the number of vehicle collisions followed the trend of the installations of the red light cameras.
+For stance, between 2016 and 2017, all four districts made the installations of lots of new red light cameras, but none of the data in Figure 4 seemed to reflect that action.
+The implication of Section \@ref(data) is that: installing more red cameras does not change the number of traffic accidents in Toronto, nor does it influence the fatality of the accidents.
+A very likely reason to explain this situation is that red light cameras only concerned the drivers who already followed the rule most of the time, but not the rule-breaking drivers.
+Another possible reason would be that at intersections with red light cameras, more drivers would intend to stop rather than to proceed when they see the light is about to turn red, resulting in a higher chance of rear end accidents [@cite2].
 
-No relationships
-No increase non decrease
-
-## Third discussion point
-
-## Limitations
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional details
+In recent years, complaints have been raised about whether red light cameras are used mainly to prevent injuries and deaths from traffic accidents or earn revenues for the government [@cite3].
+Measures should be taken to improve the effectiveness of this system and to ensure the safety of drivers and pedestrians who are the victims of the vehicle accidents.
+For instance, drivers should be more notified on the existence of red light cameras, instead of knowing them after they are fined, because red light cameras is supposed to prevent accidents but not to punish people after they cause accidents.
+Government could also seek for other solutions to reduce these accidents.
+Toronto and East York has done a great job reducing the number of motor vehicle collisions, which means other district could adopt similar traffic controls and see how it works in a few years.
 
@@
 # References
-
-https://newcanadianlife.com/why-does-toronto-have-so-much-traffic/
-
-https://www.scientificamerican.com/article/red-light-cameras-may-not-make-streets-safer/

### 2022-02-06 scripts/01-data_cleaning.R

@@ -1,14 +1,12 @@
 #### Preamble ####
-# Purpose: Clean the survey data downloaded from [...UPDATE ME!!!!!]
-# Author: Rohan Alexander [CHANGE THIS TO YOUR NAME!!!!]
-# Data: 3 January 2021
-# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
+# Purpose: Clean the survey data downloaded from Toronto Open Data Portal
+# Author: Alicia Yang
+# Data: 6 February 2021
+# Contact: aliciam.yang@utoronto.ca
 # License: MIT
 # Pre-requisites: 
 # - Need to have downloaded the ACS data and saved it to inputs/data
 # - Don't forget to gitignore it!
-# - Change these to yours
-# Any other information needed?
 
 
 #### Workspace setup ####
@@ -22,24 +20,13 @@ library(dplyr)
 # get package
 package <- show_package("0b6d3a00-7de1-440b-b47c-7252fd13929f")
 
-package
-
-# get all resources for this package
-resources <- list_package_resources("0b6d3a00-7de1-440b-b47c-7252fd13929f")
-
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
 data1 <-
   list_package_resources("0b6d3a00-7de1-440b-b47c-7252fd13929f") %>%
   get_resource()
 
 # get package
 package2 <- show_package("9fcff3e1-3737-43cf-b410-05acd615e27b")
-package2
 
-# get all resources for this package
-resources2 <- list_package_resources("9fcff3e1-3737-43cf-b410-05acd615e27b")
 
 data2 <-
   list_package_resources("9fcff3e1-3737-43cf-b410-05acd615e27b") %>%

### 2022-02-06 outputs/paper/.ipynb_checkpoints/paper-checkpoint.Rmd
discussion and introduction

--- 2022-02-06 outputs/paper/paper.Rmd
@@
 This report investigates the motor vehicle collisions in Toronto from 2006 to 2020, with a focus on the ones result in injuries or deaths.
-In Section \@ref(data), the Motor Vehicle Collisions dataset is introduced and organized to visualize the change in the annual number of traffic accidents with respect to their fatality and their location.
+In Section \@ref(motor-vehicle-collisions-involving-killed-or-seriously-injured-persons), the Motor Vehicle Collisions dataset is introduced and organized to visualize the change in the annual number of traffic accidents with respect to their fatality and their location.
 Then, I discussed about the possible implications of these visualizations.
+Figure 1 showed that there was a descending trend of traffic accidents from 2006 to 2020.
 By Figure 2, I concluded that motor vehicle collisions are always more likely to occur at intersections.
-Thus,
-Continuing
-
-You can and should cross-reference sections and sub-sections.
-For instance, Section \@ref(data).
-R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
+Thus, I brought out the Red Light Cameras dataset in Section \@ref(red-light-cameras) to see whether if the number of red light cameras has an effect on the number of traffic accidents with serious injuries and deaths.
+The result from the datasets indicates that there is no clear relationship between this two factors, suggesting that just by installing more red light cameras cannot prevent traffic accidents from happening at intersections.
+More implications and further discussions on this result is presented in Section \@ref(discussion).
 
@@
 As a result, the location of those accidents specific to neighborhoods might not be accurate.
-Therefore, I will use the variable district instead of neighbourhood to categorize the accidents later in the report to avoid this bias.
+Therefore, I will use the variable district instead of the variable neighbourhood to categorize the accidents later in the report to avoid the bias.
 
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 # Load motor vehicle collision data and clean
@@
 
-```{r fig.cap = "Comparing the number of traffic collisions each year from 2006 to 2020", fig.width=8.4,fig.height=3.6, echo=FALSE}
+```{r echo=FALSE, fig.cap="Comparing the number of traffic collisions each year from 2006 to 2020", fig.width=8.4, fig.height=3.6}
 # Create a bar plot to show number of collisions by fatality
@@
 
-```{r fig.cap = "Comparing the location of traffic collisions each year from 2006 to 2020", echo=FALSE, fig.width=6,fig.height=3.2}
+```{r echo=FALSE, fig.cap="Comparing the location of traffic collisions each year from 2006 to 2020", fig.width=6, fig.height=3.2}
 # Create a bar plot to show number of collisions by accident location
@@
 
-```{r dadaextract1, message=FALSE, warning=FALSE, echo=FALSE}
+```{r dadaextract1, echo=FALSE, message=FALSE, warning=FALSE}
 # Load red light camera data and clean
@@
 
-```{r message=FALSE, warning=FALSE, echo=FALSE, fig.cap = "Number of red light cameras in the four districts each year", fig.width=7,fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Number of red light cameras in the four districts each year", fig.width=7, fig.height=3.2}
 # Create a new dataset that presents the total number of red light cameras in each district
@@
 I wanted to see how the increase in number of red light cameras compare with the traffic collision's frequency in each district.
-So, I selected the accidents that occurred at intersections or somewhere near intersections from the Motor Vehicle Collision dataset and created a ggplot to see their trend by each district in Toronto.
+So, I selected the accidents that occurred at intersections or somewhere near intersections from the Motor Vehicle Collision dataset and created a bar plot to see their trend by each district in Toronto.
 
-```{r, echo=FALSE, warning=FALSE, fig.cap = "Comparing the number of motor vehicle collisions take place at or near intersections by district", fig.width=7,fig.height=3.2}
+```{r echo=FALSE, warning=FALSE, fig.cap="Comparing the number of motor vehicle collisions take place at or near intersections by district", fig.width=7, fig.height=3.2}
 # Select the accidents that only occur at or near intersections
@@
 The analysis of the two datasets show no relationship between the cases of motor vehicle collisions in Toronto and the number of red light cameras.
-From the comparision between Figure 3 and Figure 4, we see that in none of the districts did the number of vehicle collisions followed the trend of the installations of the red light cameras.
+From the comparison between Figure 3 and Figure 4, we see that in none of the districts did the number of vehicle collisions followed the trend of the installations of the red light cameras.
 For stance, between 2016 and 2017, all four districts made the installations of lots of new red light cameras, but none of the data in Figure 4 seemed to reflect that action.

### 2022-02-06 outputs/paper/paper.Rmd
fix3

@@
 This report focuses on four of these variables which are: year, accident location, district, and fatality.
-By using R [@citeR], and R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr], an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.
+By using R [@citeR], and R packages "tidyverse" [@citetidyverse], "dplyr" [@citedplyr], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr], an extract of the cleaned dataset (Table \@ref(tab:dataextract)) is shown below.
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 # Load motor vehicle collision data and clean
@@
 To reduce errors generated from this downside, I will only utilize the variables with no missing values.
-R [@citeR], R packages "tidyverse" [@citetidyverse], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr] are used to produce an extract of the cleaned dataset (Table \@ref(tab:dadaextract1)).
+R [@citeR], R packages "tidyverse" [@citetidyverse], "dplyr" [@citedplyr], "janitor" [@citejanitor], "kableExtra" [@citekableExtra] and "knitr" [@citeknitr] are used to produce an extract of the cleaned dataset (Table \@ref(tab:dadaextract1)).
 
-```{r dadaextract1, message=FALSE, warning=FALSE, echo=FALSE}
+```{r dadaextract1, echo=FALSE, message=FALSE, warning=FALSE}
 # Load red light camera data and clean
@@
 Government could also seek for other solutions to reduce these accidents.
-Toronto and East York has done a great job reducing the number of motor vehicle collisions, which means other district could adopt similar traffic controls and see how it works in a few years.
+Toronto and East York has done a great job reducing the number of motor vehicle collisions, which means that other district could adopt similar traffic controls and see how it works in a few years.
 

### 2022-02-06 scripts/01-data_cleaning.R

@@ -11,7 +11,6 @@
 
 #### Workspace setup ####
 # Use R Projects, not setwd().
-library(haven)
 library(tidyverse)
 library(opendatatoronto)
 library(dplyr)
@@ -27,16 +26,17 @@ data1 <-
 # get package
 package2 <- show_package("9fcff3e1-3737-43cf-b410-05acd615e27b")
 
-
 data2 <-
   list_package_resources("9fcff3e1-3737-43cf-b410-05acd615e27b") %>%
   get_resource()
 
+# Import motor vehicle collisions data
 write_csv(
   x = data1,
   file = "Telling-stories-with-data/inputs/data/vehicle_collisions.csv"
 )
 
+# Import red light cameras data
 write_csv(
   x = data2,
   file = "Telling-stories-with-data/inputs/data/red_light_camera.csv"

### 2022-02-07 outputs/paper/paper.Rmd
finalize

@@
 # Discussion
-The analysis of the two datasets show no relationship between the cases of motor vehicle collisions in Toronto and the number of red light cameras.
+The analysis of the two datasets shows no relationship between the cases of motor vehicle collisions in Toronto and the number of red light cameras.
 From the comparison between Figure 3 and Figure 4, we see that in none of the districts did the number of vehicle collisions followed the trend of the installations of the red light cameras.
-For stance, between 2016 and 2017, all four districts made the installations of lots of new red light cameras, but none of the data in Figure 4 seemed to reflect that action.
+For instance, between 2016 and 2017, all four districts made the installations of lots of new red light cameras, but none of the bar plot in Figure 4 seemed to reflect that action.
 The implication of Section \@ref(data) is that: installing more red cameras does not change the number of traffic accidents in Toronto, nor does it influence the fatality of the accidents.
@@
 In recent years, complaints have been raised about whether red light cameras are used mainly to prevent injuries and deaths from traffic accidents or earn revenues for the government [@cite3].
-Measures should be taken to improve the effectiveness of this system and to ensure the safety of drivers and pedestrians who are the victims of the vehicle accidents.
-For instance, drivers should be more notified on the existence of red light cameras, instead of knowing them after they are fined, because red light cameras is supposed to prevent accidents but not to punish people after they cause accidents.
-Government could also seek for other solutions to reduce these accidents.
-Toronto and East York has done a great job reducing the number of motor vehicle collisions, which means that other district could adopt similar traffic controls and see how it works in a few years.
+Measures should be taken to improve the effectiveness of this system and to ensure the safety of the drivers and pedestrians who are victims of the vehicle accidents.
+For instance, drivers should be more notified on the existence of red light cameras instead of knowing that after they are fined, because red light cameras are supposed to prevent severe accidents from happening but not to punish people after they cause the accidents.
+Government could also seek for other solutions to reduce those vehicle collisions.
+Toronto and East York has done a great job reducing the number of traffic collisions from 2006 to 2020, which means that other district could adopt similar traffic controls and see how it works in a few years.
 
